---
title: "EDA Predictive Maintenance dataset1"
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'EDAPredictiveMaintenancedataset1.html'))})
author: "Sridhar Kothalanka"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Overview
In the below exercise EDA is performed on a predictive maintenance data set to understand the data

## Steps
1. Load the necessary R libraries
2. Load the average utilization summary data
3. Plot the average utilization summary data
4. Quick observations
5. A closer look at each of the 2 stations
6. Station-1 how does the utilization look over a period of 1 year
7. Quick observations on Station-1
8. Station-2 how does the utilization look over a period of 1 year
9. Quick observations on Station-2
10. Station-1 and Station-2 side-by-side comparison

### 1. Load the necessary R libraries
```{r results="hide", message=FALSE, warning=FALSE, error=FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)
library(tidyr)
```

### 2. Load the average utilization summary data

The raw data sets consists of observations of two stations across 3 shifts and its utilization over a period of 1 year. The raw data was processed and features like "day of the week" was added to get a quick understanding of any plant behavior patterns.

```{r tidy=TRUE}
df1 <- read.csv("../data/eda-pm-ds1/station1n2-avgUtilization.csv")
names(df1)[2]<-"avgUtilization"
```

### 3. Plot the average utilization summary data

Plotting the average utilization of all stations for each day of the week help us to understand if there are any weekly cycles

```{r tidy=TRUE}
annotation  <- data.frame(x=c(6,7),y=c(17,5),label=c("reduced shifts?","holiday?"))
df1 %>% 
  ggplot(aes(x=factor(df1$weekDayName,levels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")), y=avgUtilization)) + 
  geom_bar(stat="identity", width=0.5) +
  xlab("Day of week") +
  ylab("Average Utilization % \n(2 Stations and across 3 shifts)") +
  geom_text(data=annotation, aes(x=x,y=y,label=label, color="red", size=7, angle=0, fontface="bold")) +
  theme(legend.position = "none") +
  ggtitle("Average Utilization per shift on each day of the week")

```

### 4. Quick observations
What we can infer from the above plot is:

1. On a Sunday there is little or no activity across the shifts, possibly it may be holiday
2. On a Saturday there is some reduction in shifts as a result the average is lower then the rest of the week days

### 5. A closer look at each of the 2 stations
Let us take a closer look at each of the two stations independently

* Station-1
* Station-2

and see how things show up

### 6. Station-1 how does the utilization look over a period of 1 year?
```{r tidy=TRUE}
df2 <- read.csv("../data/eda-pm-ds1/station1.csv") %>%
  gather(key="Shift",value="utilization","Shift1","Shift2","Shift3")
df2$Shift <- factor(df2$Shift,levels=c("Shift1","Shift2","Shift3"))

dfx <- df2 %>% filter(df2$Shift == "Shift1" & df2$weekDayName != "Sun") %>% select(utilization)
shift1AvgExSun <- mean(dfx$utilization, na.rm = TRUE)
dfx <- df2 %>% filter(df2$Shift == "Shift2" & df2$weekDayName != "Sun") %>% select(utilization)
shift2AvgExSun <- mean(dfx$utilization, na.rm = TRUE)
dfx <- df2 %>% filter(df2$Shift == "Shift3" & df2$weekDayName != "Sun") %>% select(utilization)
shift3AvgExSun <- mean(dfx$utilization, na.rm = TRUE)

df2 %>% 
  ggplot(aes(x=factor(df2$weekDayName,levels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")), y=utilization,fill=Shift)) + 
  geom_bar(position="dodge",stat="identity",width=0.6) +
  ggtitle("Station1") +
  ylab("Utilization %") +
  xlab("Day of week") +
  ylim(0,40) +
  scale_fill_manual("legend", values = c("Shift1" = "#00BFC4", "Shift2" = "#7CAE00", "Shift3" = "#F8766D")) +
  ggtitle("Station-1 utilization breakdown") +
  theme(legend.title=element_blank()) +
  annotate(geom="curve",x=6.5,y=8,xend=7.1, yend=5, curvature=0.3, arrow=arrow(length=unit(2,"mm"))) +
  annotate(geom="text",x=6.25,y=12,label="shift3 utilization \ngoes up before \nworkweek",hjust="left") +
  geom_hline(yintercept=shift1AvgExSun, linetype="dashed", color = "#00BFC4", size=1) +
  geom_hline(yintercept=shift2AvgExSun, linetype="dashed", color = "#7CAE00", size=1) +
  geom_hline(yintercept=shift3AvgExSun, linetype="dashed", color = "#F8766D", size=1) +
  annotate(geom="text",x=6.75,y=shift1AvgExSun+1,label="avg (M-F)",hjust="left")
  
```

### 7. Quick observations on Station-1
1. Utilization is below 100%
2. Shift 3 seems to be contributing more towards the decrease in utilization
3. As before Saturday is running reduced shifts and Sunday may be a holiday


### 8. Station-2 how does the utilization look over a period of 1 year?
```{r tidy=TRUE}
df3 <- read.csv("../data/eda-pm-ds1/station2.csv") %>%
  gather(key="Shift",value="utilization","Shift1","Shift2","Shift3")
df3$Shift <- factor(df3$Shift,levels=c("Shift1","Shift2","Shift3"))

dfx <- df3 %>% filter(df3$Shift == "Shift1" & df3$weekDayName != "Sun") %>% select(utilization)
shift1AvgExSun <- mean(dfx$utilization, na.rm = TRUE)
dfx <- df3 %>% filter(df3$Shift == "Shift2" & df3$weekDayName != "Sun") %>% select(utilization)
shift2AvgExSun <- mean(dfx$utilization, na.rm = TRUE)
dfx <- df3 %>% filter(df3$Shift == "Shift3" & df3$weekDayName != "Sun") %>% select(utilization)
shift3AvgExSun <- mean(dfx$utilization, na.rm = TRUE)

df3 %>% 
  ggplot(aes(x=factor(df3$weekDayName,levels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")), y=utilization,fill=Shift)) + 
  geom_bar(position="dodge",stat="identity",width=0.6) +
  ggtitle("Station1") +
  ylab("Utilization %") +
  xlab("Day of week") +
  ylim(0,40) +
  scale_fill_manual("legend", values = c("Shift1" = "#00BFC4", "Shift2" = "#7CAE00", "Shift3" = "#F8766D")) +
  ggtitle("Station-2 utilization breakdown") +
  theme(legend.title=element_blank()) +
  annotate(geom="curve",x=6.5,y=12,xend=7.1, yend=10, curvature=0.3, arrow=arrow(length=unit(2,"mm"))) +
  annotate(geom="text",x=6.25,y=16,label="shift3 utilization \ngoes up before \nworkweek",hjust="left") +
  geom_hline(yintercept=shift1AvgExSun, linetype="dashed", color = "#00BFC4", size=1) +
  geom_hline(yintercept=shift2AvgExSun, linetype="dashed", color = "#7CAE00", size=1) +
  geom_hline(yintercept=shift3AvgExSun, linetype="dashed", color = "#F8766D", size=1) +
  annotate(geom="text",x=6.75,y=shift1AvgExSun+1,label="avg (M-F)",hjust="left")
```

### 9. Quick observations on Station-2
1. Utilization is below 100%
2. Shift 3 seems to be contributing more towards the decrease in utilization
3. As before Saturday is running reduced shifts and Sunday may be a holiday

### 10. Station-1 and Station-2 side-by-side comparison

```{r tidy=TRUE}
df4 <- read.csv("../data/eda-pm-ds1/station1n2.csv") 
df4$Station_ID[df4$Station_ID == 1] <- "Station-1"
df4$Station_ID[df4$Station_ID == 2] <- "Station-2"
df4$Shift[df4$Shift == 1] <- "Shift1"
df4$Shift[df4$Shift == 2] <- "Shift2"
df4$Shift[df4$Shift == 3] <- "Shift3"

df4 %>% 
  ggplot(aes(x=factor(df4$weekDayName,levels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")), y=avgUtilization,fill=Shift)) + 
  geom_bar(position="dodge",stat="identity") +
  ylab("Utilization %") +
  xlab("Day of week") +
  scale_fill_manual("legend", values = c("Shift1" = "#00BFC4", "Shift2" = "#7CAE00", "Shift3" = "#F8766D")) +
  ggtitle("Station-1 and Station-2 side-by-side comparison") +
  theme(legend.title=element_blank()) +
  facet_grid(~Station_ID) 
```  


