---
title: "EDA Predictive Maintenance Dataset2"
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'EDAPredictiveMaintenancedataset2.html'))})
author: "Sridhar Kothalanka"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
In the below exercise EDA is performed on a predictive maintenance data set to understand the data "hello"


### 1. Load the necessary R libraries
```{r results="hide", message=FALSE, warning=FALSE, error=FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)
library(tidyr)
```

### 2. Load the average utilization summary data

The raw data sets consists of observations of two stations across 3 shifts and its utilization over a period of 1 year. The raw data was processed and features like "day of the week" was added to get a quick understanding of any plant behavior patterns.

```{r tidy=TRUE}
df1 <- read.csv("../data/eda-pm-ds1/station1n2-avgUtilization.csv")
names(df1)[2]<-"avgUtilization"
```
